Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-01-17 ms/manuscript.Rmd
Make lintr happy (ms)

# Set knitr options
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)

@@

```{r intro-stats}
# tally number of non-hybrid pteridophytes with known repro. mode
repro_count <-
repro_data %>%
  count(reproductive_mode) %>%
  mutate(mode = if_else(reproductive_mode == "unknown", "unknown", "known")) %>%
  group_by(mode) %>%
  summarize(total = sum(n)) %>%
  split(.$mode)

repro_count$known$percent <- (repro_count$known$total/(repro_count$unknown$total + repro_count$known$total)) %>% percent(accuracy = 0.1)
```

@@

# Verify that tips in the plastome tree are accounted for
# by FTOL sampling + japan fern rbcL
tibble(taxon = plastome_tree$tip.label) %>%
  anti_join(select(plastome_sampling, taxon) %>% unique, by = "taxon") %>%
  anti_join(tibble(taxon = rownames(japan_rbcL)), by = "taxon") %>%
  verify(nrow(.) == 0, success_fun = success_logical)

@@

```{r dataset-stats}
### Taxon list ###
# Make list of all native ferns in Japan (based off Green List), *including* hybrids
fern_list <-
  green_list %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  left_join(ppgi, by = "genus") %>%
  filter(class == "Polypodiopsida")

fern_list_non_hybrid <-
  fern_list %>%
  filter(hybrid == FALSE) %>%
  mutate(gn_parse_tidy(.$scientific_name) %>% select(taxon = canonicalsimple)) %>%
  mutate(taxon = str_replace_all(taxon, " ", "_"))

@@

# Check height (age) of fern phylogeny
phylogeny_age <- japan_fern_tree %>% phytools::nodeHeights() %>% max %>% round(1)

# Check monophyly of genera
mono_status_genera <-
  tibble(
    tip = japan_fern_tree$tip.label,
    tip_no = 1:length(tip),
    genus = str_split(tip, "_") %>% map_chr(1)
  ) %>%
  group_by(genus) %>%
  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))

@@

# Check monophyly of families
mono_status_families <-
  tibble(
    tip = japan_fern_tree$tip.label,
    tip_no = 1:length(tip),
    genus = str_split(tip, "_") %>% map_chr(1)
  ) %>%
  left_join(ppgi, by = "genus") %>%
  group_by(family) %>%
  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))

@@

# Check monophyly of ferns
mono_status_class <- 
  tibble(
    tip = japan_pterido_tree$tip.label,
    tip_no = 1:length(tip),
    genus = str_split(tip, "_") %>% map_chr(1)
  ) %>%
  left_join(ppgi, by = "genus") %>%
  group_by(class) %>%
  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_pterido_tree, tips = .)))

@@

# Count number of traits of each type in final dataset
n_binary_traits <- traits_summary %>% filter(trait_type == "binary") %>% nrow() %>% jntools::english2()
n_cont_traits <- traits_summary %>% filter(trait_type == "continuous") %>% nrow() %>% jntools::english2()
n_qual_traits <- traits_summary %>% filter(trait_type == "qualitative") %>% nrow() %>% jntools::english2()
n_qual_traits_as_binary <- traits_summary %>% filter(trait_type == "qualitative") %>% pull(n) %>% sum %>% jntools::english2()

@@

n_taxa_traits_sampled <- sum(trait_sampling_summary$sampled)
percent_taxa_traits_sampled <- (n_taxa_traits_sampled / total_fern_taxa) %>% scales::percent(accuracy = 0.1)

@@

```{r discus-model-stats}
# Check that conditional AIC was lower in the reproductive model compared to 
# the environmental model for SES of RPD
# Split aic_env_repro dataframe into nested list for easier indexing
c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
assert_that(c_aic$rpd_obs_z$percent_apo$conditional_aic < c_aic$rpd_obs_z$temp$conditional_aic)
# But not for SES of PD
assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic > c_aic$pd_obs_z$temp$conditional_aic)
```

@@

```{r conserv-stats}
# Convert signif_cells_protected_area to wide format
signif_cells_protected_area_wide <-
  signif_cells_protected_area %>% select(status, metric, percent_protected) %>%
  pivot_wider(names_from = status, values_from = percent_protected) %>%
  rowwise %>%
  mutate(total = sum(medium, high)) %>%
  ungroup

# Calculate min and max percent area protected overall
min_protected_med <- signif_cells_protected_area_wide %>% pull(medium) %>% min %>% scales::percent(accuracy = 0.1)
max_protected_med <- signif_cells_protected_area_wide %>% pull(medium) %>% max %>% scales::percent(accuracy = 0.1)
min_protected_high <- signif_cells_protected_area_wide %>% pull(high) %>% min %>% scales::percent(accuracy = 0.1)
max_protected_high <- signif_cells_protected_area_wide %>% pull(high) %>% max %>% scales::percent(accuracy = 0.1)
min_protected_total <- signif_cells_protected_area_wide %>% pull(total) %>% min %>% scales::percent(accuracy = 0.1)
max_protected_total <- signif_cells_protected_area_wide %>% pull(total) %>% max %>% scales::percent(accuracy = 0.1)

# Calculate min and max protected area for the best protected category
signif_cells_protected_area_best <-
  signif_cells_protected_area_wide %>%
  slice_max(total)

@@

# Calculate the total area of Japan
area_of_japan <-
  japan_shp %>%
  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
  pull(area)

# Calculate baseline protection percentage: percent protected areas across the whole country
total_percent_protected <-
  protected_areas %>%
  st_make_valid %>%
  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  filter(status %in% c("medium", "high")) %>%
  group_by(status) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_protected = total_area/area_of_japan) %>%
  mutate(percent_protected = as.numeric(percent_protected)) %>%
  # Add sum for high + medium
  bind_rows(
    summarize(
      ., 
      percent_protected = sum(percent_protected),
      total_area = sum(total_area)
    ) %>% mutate(status = "total")
  )

# Extract values for medium and high protection level
total_percent_protected_medium <- total_percent_protected %>% filter(status == "medium") %>% pull(percent_protected)
total_percent_protected_high <- total_percent_protected %>% filter(status == "high") %>% pull(percent_protected)
total_percent_protected_all <- sum(total_percent_protected_medium, total_percent_protected_high)

@@

# Calculate baseline protection percentage: percent protected areas across the whole country
total_percent_deer_danger <-
  deer_range %>%
  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  group_by(range) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_danger = total_area/area_of_japan) %>%
  mutate(percent_danger = as.numeric(percent_danger))

# Convert to a list for easy indexing
deer_overall <- total_percent_deer_danger %>%
  mutate(
    # Format numbers for nice printing
    area = total_area %>% as.numeric %>% scales::number(accuracy = 1, big.mark = ",") %>% paste("km^2^"),
    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
  split(.$range)

@@

n_southern_genera <- southern_genera %>% pull(higher_taxon) %>% n_distinct %>% english2

@@

n_southern_families <- southern_families %>% pull(higher_taxon) %>% n_distinct %>% english2
```

@@

# Make map
japan_map_plot <- ggplot(japan) + 
  geom_sf(size = 0.25) +
  geom_segment(data = japan_lines, aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)) +
  scale_linetype_manual(values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")) +
  geom_richtext(
    data = japan_points, 
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  annotation_scale(location = "br", width_hint = 0.2) +
  map_theme() +
  theme(legend.position = "none")
```

```{r japan-map-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("japan-map"), "pdf"), japan_map_plot, width = 5, height = 5, units = "in")
```

@@

```{r raw-diversity-make}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

# Raw PD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(axis.text.y = element_blank())

# Raw FD
c <- ggplot() + 
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
  map_theme() +
  theme(axis.text.y = element_blank())

@@

```{r raw-diversity-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7.25, height = 4, units = "in", raw_diversity_plot)
```

@@

```{r rand-diversity-make}
# PD
a <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()
  )

# RPD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

# FD
c <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  map_theme() +
  theme(legend.title = element_blank())

# RFD
d <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**D** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  annotation_scale(location = "br", width_hint = 0.2) +
  map_theme() +
  theme(
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

rand_diversity_plot <- a + b + c + d + 
  plot_layout(ncol = 2, nrow = 2, guides = 'collect') &
  theme(legend.position = "bottom")
```

```{r rand-diversity-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7.25, height = 8, units = "in")
```

@@

# Format labels for printing
resp_var_lookup <-
  tibble(
    resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
    resp_var_print = c("Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
  )

@@

  ggplot(plot_data, aes(x = t_value, y = term_print, fill = signif)) +
    geom_vline(xintercept = 0, linetype = "dotted") +
    geom_col() +
    # scale_x_continuous(n.breaks = 4) +
    scale_x_continuous() +
    scale_fill_manual(
      values = c("TRUE" = "grey20", "FALSE" = "grey70"),
      labels = c("TRUE" = "*p* < 0.05", "FALSE" = "*p* > 0.05")) +
    labs(
      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"),
      fill = "Significance") +
    theme_bw(base_size = 9) +
    theme(
      axis.title = element_blank(),
      axis.text.y = element_markdown(),
      legend.text = element_markdown(),
      panel.grid.minor = element_blank(),
      legend.position = "none")
}

# Loop across data and make a plot of each
coeff_subplots <-
  model_params_for_plot %>%
  group_by(resp_var, model_type) %>%
  nest() %>%
  ungroup %>%
  left_join(model_subplot_order, by = c("resp_var", model_type = "indep_var")) %>%
  arrange(order) %>%
  mutate(plot = map(data, plot_coeff)) %>%
  pull(plot)

@@

```{r coeff-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("coeff"), "pdf"), coeff_plot, width = 7.25, height = 5, units = "in")
```

@@

```{r modelfit-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("modelfit"), "pdf"), modelfit_plot, width = 7.25, height = 5, units = "in")
```

@@

```{r bioregions-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7.25, height = 5, units = "in")
```

>>>>

# Set knitr options
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "hide", cache = FALSE)

@@

```{r intro-stats}
# Tally number of non-hybrid pteridophytes with known repro. mode
repro_count <-
repro_data %>%
  count(reproductive_mode) %>%
  mutate(mode = if_else(reproductive_mode == "unknown", "unknown", "known")) %>%
  group_by(mode) %>%
  summarize(total = sum(n)) %>%
  split(.$mode)

# Calculate percent of species with known reproductive mode
repro_count$known$percent <-
  repro_count$known$total %>%
  magrittr::divide_by(repro_count$unknown$total + repro_count$known$total) %>%
  percent(accuracy = 0.1)
```

@@

# Verify that tips in the plastome tree are accounted for
# by FTOL sampling + japan fern rbcL
tibble(taxon = plastome_tree$tip.label) %>%
  anti_join(select(plastome_sampling, taxon) %>% unique, by = "taxon") %>% # nolint
  anti_join(tibble(taxon = rownames(japan_rbcL)), by = "taxon") %>%
  verify(nrow(.) == 0, success_fun = success_logical)

@@

```{r dataset-stats}
### Taxon list ###
# Make list of all native ferns in Japan (based off Green List)
# *including* hybrids
fern_list <-
  green_list %>%
  mutate(
    genus = str_split(taxon, "_") %>%
      map_chr(1)
  ) %>%
  left_join(ppgi, by = "genus") %>%
  filter(class == "Polypodiopsida")

fern_list_non_hybrid <-
  fern_list %>%
  filter(hybrid == FALSE) %>%
  mutate(
    gn_parse_tidy(.$scientific_name) %>%
      select(taxon = canonicalsimple)
  ) %>%
  mutate(taxon = str_replace_all(taxon, " ", "_"))

@@

# Check height (age) of fern phylogeny
phylogeny_age <- japan_fern_tree %>%
  phytools::nodeHeights() %>%
  max() %>%
  round(1)

# Check monophyly of genera
mono_status_genera <-
  tibble(
    tip = japan_fern_tree$tip.label,
    tip_no = seq_along(tip),
    genus = str_split(tip, "_") %>% map_chr(1)
  ) %>%
  group_by(genus) %>%
  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
  mutate(is_mono = map_lgl(
    tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))

@@

# Check monophyly of families
mono_status_families <-
  tibble(
    tip = japan_fern_tree$tip.label,
    tip_no = seq_along(tip),
    genus = str_split(tip, "_") %>% map_chr(1)
  ) %>%
  left_join(ppgi, by = "genus") %>%
  group_by(family) %>%
  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
  mutate(is_mono = map_lgl(
    tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))

@@

# Check monophyly of ferns
mono_status_class <-
  tibble(
    tip = japan_pterido_tree$tip.label,
    tip_no = seq_along(tip),
    genus = str_split(tip, "_") %>% map_chr(1)
  ) %>%
  left_join(ppgi, by = "genus") %>%
  group_by(class) %>%
  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
  mutate(is_mono = map_lgl(
    tip_vec, ~ape::is.monophyletic(phy = japan_pterido_tree, tips = .)))

@@

# Count number of traits of each type in final dataset
n_binary_traits <- traits_summary %>%
  filter(trait_type == "binary") %>%
  nrow() %>%
  jntools::english2()
n_cont_traits <- traits_summary %>%
  filter(trait_type == "continuous") %>%
  nrow() %>%
  jntools::english2()
n_qual_traits <- traits_summary %>%
  filter(trait_type == "qualitative") %>%
  nrow() %>%
  jntools::english2()
n_qual_traits_as_binary <- traits_summary %>%
  filter(trait_type == "qualitative") %>%
  pull(n) %>%
  sum() %>%
  jntools::english2()

@@

n_taxa_traits_sampled <- sum(trait_sampling_summary$sampled)

percent_taxa_traits_sampled <-
  n_taxa_traits_sampled %>%
  magrittr::divide_by(total_fern_taxa) %>%
  scales::percent(accuracy = 0.1)

@@

```{r discus-model-stats}
# Check that conditional AIC was lower in the reproductive model compared to
# the environmental model for SES of RPD
# Split aic_env_repro dataframe into nested list for easier indexing
c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
assert_that(
  c_aic$rpd_obs_z$percent_apo$conditional_aic <
    c_aic$rpd_obs_z$temp$conditional_aic)
# But not for SES of PD
assert_that(
  c_aic$pd_obs_z$percent_apo$conditional_aic >
    c_aic$pd_obs_z$temp$conditional_aic)
```

@@

```{r conserv-stats}
# Convert signif_cells_protected_area to wide format
signif_cells_protected_area_wide <- #nolint
  signif_cells_protected_area %>%
  select(status, metric, percent_protected) %>%
  pivot_wider(names_from = status, values_from = percent_protected) %>%
  rowwise %>%
  mutate(total = sum(medium, high)) %>%
  ungroup

# Calculate min and max percent area protected overall
# - define helper functions
calc_min_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
  signif_cells_protected_area_wide %>%
    pull(...) %>%
    min() %>%
    scales::percent(accuracy = 0.1)
}
calc_max_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
  signif_cells_protected_area_wide %>%
    pull(...) %>%
    max() %>%
    scales::percent(accuracy = 0.1)
}

min_protected_med <- calc_min_pa(signif_cells_protected_area_wide, medium)
max_protected_med <- calc_max_pa(signif_cells_protected_area_wide, medium)
min_protected_high <- calc_min_pa(signif_cells_protected_area_wide, high)
max_protected_high <- calc_max_pa(signif_cells_protected_area_wide, high)
min_protected_total <- calc_min_pa(signif_cells_protected_area_wide, total)
max_protected_total <- calc_max_pa(signif_cells_protected_area_wide, total)

# Calculate min and max protected area for the best protected category
signif_cells_protected_area_best <- # nolint
  signif_cells_protected_area_wide %>%
  slice_max(total)

@@

# Calculate the total area of Japan
area_of_japan <-
  japan_shp %>%
  mutate(
    area = st_area(.) %>%
      units::set_units(km^2)
  ) %>%
  pull(area)

# Calculate baseline protection percentage:
# percent protected areas across the whole country
total_percent_protected <-
  protected_areas %>%
  st_make_valid %>%
  mutate(
    area = st_area(.) %>%
      units::set_units(km^2)
  ) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  filter(status %in% c("medium", "high")) %>%
  group_by(status) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_protected = total_area / area_of_japan) %>%
  mutate(percent_protected = as.numeric(percent_protected)) %>%
  # Add sum for high + medium
  bind_rows(
    summarize(
      .,
      percent_protected = sum(percent_protected),
      total_area = sum(total_area)
    ) %>% mutate(status = "total")
  )

# Extract values for medium and high protection level
total_percent_protected_medium <- total_percent_protected %>%
  filter(status == "medium") %>%
  pull(percent_protected)
total_percent_protected_high <- total_percent_protected %>%
  filter(status == "high") %>%
  pull(percent_protected)
total_percent_protected_all <- sum(
  total_percent_protected_medium, total_percent_protected_high)

@@

# Calculate baseline protection percentage:
# percent protected areas across the whole country
total_percent_deer_danger <-
  deer_range %>%
  mutate(
    area = st_area(.) %>% units::set_units(km^2)
  ) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  group_by(range) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_danger = total_area / area_of_japan) %>%
  mutate(percent_danger = as.numeric(percent_danger))

# Convert to a list for easy indexing
deer_overall <- total_percent_deer_danger %>%
  mutate(
    # Format numbers for nice printing
    area = total_area %>%
      as.numeric %>%
      scales::number(accuracy = 1, big.mark = ",") %>%
      paste("km^2^"),
    percent = scales::percent(percent_danger, accuracy = 0.1)
  ) %>%
  split(.$range)

@@

n_southern_genera <- southern_genera %>%
  pull(higher_taxon) %>%
  n_distinct() %>%
  english2()

@@

n_southern_families <- southern_families %>%
  pull(higher_taxon) %>%
  n_distinct() %>%
  english2()
```

@@

# Make map
japan_map_plot <- ggplot(japan) +
  geom_sf(size = 0.25) +
  geom_segment(
    data = japan_lines,
    aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)
  ) +
  scale_linetype_manual(
    values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")
  ) +
  geom_richtext(
    data = japan_points,
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  annotation_scale(location = "br", width_hint = 0.2) +
  map_theme() +
  theme(legend.position = "none")
```

```{r japan-map-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("japan-map"), "pdf"),
  japan_map_plot, width = 5, height = 5, units = "in")
```

@@

```{r raw-diversity-make}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**A** Richness</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") +
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

# Raw PD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**B** PD</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") +
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(axis.text.y = element_blank())

# Raw FD
c <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**C** FD</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") +
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
  map_theme() +
  theme(axis.text.y = element_blank())

@@

```{r raw-diversity-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("raw-diversity"), "pdf"),
  width = 7.25, height = 4, units = "in", raw_diversity_plot)
```

@@

```{r rand-diversity-make}
# PD
a <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") + #nolint
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**A** PD</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()
  )

# RPD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**B** RPD</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

# FD
c <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**C** FD</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  map_theme() +
  theme(legend.title = element_blank())

# RFD
d <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  annotate(
    geom = "richtext", y = Inf, x = -Inf,
    label = "<span style='color:black'>**D** RFD</span>",
    hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  annotation_scale(location = "br", width_hint = 0.2) +
  map_theme() +
  theme(
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

rand_diversity_plot <- a + b + c + d +
  plot_layout(ncol = 2, nrow = 2, guides = "collect") &
  theme(legend.position = "bottom")
```

```{r rand-diversity-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("rand-diversity"), "pdf"),
  rand_diversity_plot, width = 7.25, height = 8, units = "in")
```

@@

# Format labels for printing
resp_var_lookup <-
  tibble(
    resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
    resp_var_print = c(
      "Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
  )

@@

  ggplot(plot_data, aes(x = t_value, y = term_print, fill = signif)) +
    geom_vline(xintercept = 0, linetype = "dotted") +
    geom_col() +
    # scale_x_continuous(n.breaks = 4) +
    scale_x_continuous() +
    scale_fill_manual(
      values = c("TRUE" = "grey20", "FALSE" = "grey70"),
      labels = c("TRUE" = "*p* < 0.05", "FALSE" = "*p* > 0.05")) +
    labs(
      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"), #nolint
      fill = "Significance") +
    theme_bw(base_size = 9) +
    theme(
      axis.title = element_blank(),
      axis.text.y = element_markdown(),
      legend.text = element_markdown(),
      panel.grid.minor = element_blank(),
      legend.position = "none")
}

# Loop across data and make a plot of each
coeff_subplots <-
  model_params_for_plot %>%
  group_by(resp_var, model_type) %>%
  nest() %>%
  ungroup %>%
  left_join(
    model_subplot_order, by = c("resp_var", model_type = "indep_var")) %>%
  arrange(order) %>%
  mutate(plot = map(data, plot_coeff)) %>%
  pull(plot)

@@

```{r coeff-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("coeff"), "pdf"),
  coeff_plot, width = 7.25, height = 5, units = "in")
```

@@

```{r modelfit-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("modelfit"), "pdf"),
  modelfit_plot, width = 7.25, height = 5, units = "in")
```

@@

```{r bioregions-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("bioregions"), "pdf"),
  bioregions_plot, width = 7.25, height = 5, units = "in")
```
